RDF Constraint Types to Validate Metadata \\ on Highly-Complex Person-Level and Aggregated Data

### 2015-02-10 paper/paper 2/paper 2.tex
running example

There is no constraint language which can be seen as the standard.
Although, there are multiple constraint languages (having different syntaxes and semantics) which can be used to express RDF constraints (such as existential and universal quantification, cardinality restrictions, and exclusive or of properties).
The five most promising ones on being the standard are
Description Set Profiles (DSP)\footnote{\url{http://dublincore.org/documents/2008/03/31/dc-dsp/}},
Resource Shapes (ReSh)\footnote{\url{http://www.w3.org/Submission/shapes/}}, 
Shape Expressions (ShEx)\footnote{\url{http://www.w3.org/Submission/shex-primer/}},
the SPARQL Inferencing Notation (SPIN)\footnote{\url{http://spinrdf.org/}}, 
and the Web Ontology Language (OWL 2)\footnote{\url{http://www.w3.org/TR/owl2-syntax/}}.
%\section{Ideas}
%constraint design patterns:
% - ontology design pattern.org
% - DSP uses other design pattern
% - constraint elements are the same / 
% - constraint / constraint elements / constraint design patterns / constraint language
%dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
%min car more powerful than req. propery
The idea behind this paper is to express any constraint type (corresponding to a requirement to formulate RDF constraints) in a generic way using a lightweight ontology (consisting of only a few terms).

@@

\section{include this in next section}

With minimum qualified cardinality restrictions, researchers from the library domain can state
that \emph{publications} must have at least one \emph{author} which must be a \emph{person}.
This constraint can be represented generically in DL as follows: \ms{Publication $\sqsubseteq$ $\geq$1 author.Person}.

The same constraint may also be represented specifically by multiple constraint languages like OWL 2, ShEx, ReSh, or DSP:

@@

As \ms{Janeway} commands the vessel \ms{Voyager} (\ms{commandsVessel(Janeway,Voyager)}) and \ms{Voyager} is a \ms{Vessel} (\ms{rdf:type(Voyager,Vessel)}), the assignment of \ms{Janeway} to the class \ms{StarfleetCaptain} (\ms{rdf:type(Janeway,StarfleetCaptain)}) does not cause a constraint violation.
A constraint violation is raised, however, when there is not at least one \ms{commandsVessel} relationship with \ms{Janeway} as the subject. 
A constraint violation is also raised, when there is a triple \ms{commandsVessel(Janeway,Voyager)}, but \ms{Voyager} is not explicitly defined to be a \ms{Vessel}, 
i.e. the triple \ms{rdf:type(Voyager,Vessel)} is not present in the data. 
As \ms{DS9 is not a} \ms{Vessel}, 
the triples \ms{commandsVessel(Sisko,DS9)}, \ms{rdf:type(Sisko,StarfleetCaptain)}, and \ms{rdf:type(DS9,SpaceStation)}  
also cause a constraint violation.

The following table shows the mapping between the cardinality restriction expressed in DL 
and the cardinality restriction expressed by our developed ontology to describe any RDF constraint generically (see section~\ref{sec:ontology}):

\begin{gcotable}
property & StarfleetCaptain & commandsVessel & - & Vessel & $\geq$ & 1 \\
\end{gcotable}

According to this ontology, each constraint is either a constraint on classes or a constraint on properties.
The generically expressed cardinality restriction is represented in RDF as follows:

\begin{ex}
[   a PropertyConstraint ;
    contextClass StarfleetCaptain ;
    leftProperties ( commandsVessel ) ;
    classes ( Vessel ) ;
    constrainingElement ">=" ;
    constrainingValue 1 ] .
\end{ex}

-----

The generically expressed cardinality restriction is validated within our validation environment (see section \ref{sec:validation}) by means of a SPARQL query:
\begin{ex}
[   a PropertyConstraint ;
    contextClass ?cc ;
    leftProperties ( ?p1 ) ;
    classes ( ?c1 ) ;
    constrainingElement ">=" ;
    constrainingValue ?cv ] .
?subject rdf:type ?cc .
BIND ( qualifiedCardinality( ?subject, ?p1, ?c1 ) AS ?c ) .
FILTER ( ?c < ?cv ) .		  
\end{ex}

\section{Ontology to Describe RDF Constraints Generically} 
\label{sec:ontology}

In order to define an ontology to describe RDF constraints generically, it is needed to define the terminology for the formulation of RDF constraints and to classify them. 
A \ms{constraint language} is a language which is used to formulate constraints.
The W3C Data Shapes working group defines \ms{constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
We identified four dimensions to classify constraints:
\begin{itemize}
  \item \ms{Universality:} specific constraints vs. generic constraints
	\item \ms{Complexity:} simple constraints vs. complex constraints
	\item \ms{Context:} property constraints vs. class constraints
	\item \ms{DL Expressivity:} constraints expressible in DL vs. constraints not expressible in DL
\end{itemize}

As there are already five promising constraint languages, our purpose is not to invent a new constraint language.
We rather developed a very simple ontology (only three classes, three object properties, and three data properties) which is universal enough to describe any RDF constraint expressible by any RDF constraint language (see the conceptual model in figure \ref{fig:RDF-CO-conceptual-model}).
We call this ontology the \ms{RDF constraints ontology (RDF-CO)}\footnote{Available at: \url{https://github.com/boschthomas/RDF-constraints-ontology}}.

>>>>

There is no constraint language which can be seen as the standard.
Although, there are multiple constraint languages (having different syntaxes and semantics) which can be used to express RDF constraints (such as existential and universal quantification, cardinality restrictions, and exclusive or of properties).
The five most promising ones on being the standard are
Description Set Profiles (DSP)\footnote{\url{http://dublincore.org/documents/2008/03/31/dc-dsp/}},
Resource Shapes (ReSh)\footnote{\url{http://www.w3.org/Submission/shapes/}}, 
Shape Expressions (ShEx)\footnote{\url{http://www.w3.org/Submission/shex-primer/}},
the SPARQL Inferencing Notation (SPIN)\footnote{\url{http://spinrdf.org/}}, 
and the Web Ontology Language (OWL 2)\footnote{\url{http://www.w3.org/TR/owl2-syntax/}}.
%\section{Ideas}
%constraint design patterns:
% - ontology design pattern.org
% - DSP uses other design pattern
% - constraint elements are the same / 
% - constraint / constraint elements / constraint design patterns / constraint language
%dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
%min car more powerful than req. propery
The idea behind this paper is to express any constraint type (corresponding to a requirement to formulate RDF constraints) in a generic way using a lightweight vocabulary (consisting of only a few terms).

@@

\section{Running Example}

With minimum qualified cardinality restrictions, \emph{publications} can be enforced to have at least one \emph{author} which must be a \emph{person}.
This constraint can be represented generically (\emph{generic constraint}) using DL (\ms{Publication $\sqsubseteq$ $\geq$1 author.Person}), but also specifically (\emph{specific constraint}) by multiple constraint languages like OWL 2, ShEx, ReSh, or DSP:

@@

As each \emph{publication} must have at least one \emph{author} relationship to a \emph{person} 
and as the book \emph{The-Lord-Of-The-Rings} is a \emph{publication} \ms{rdf:type(The-Lord-Of-The-Rings, Publication)}, 
\emph{The-Lord-Of-The-Rings} must have at least one \emph{author} relationship to a \emph{person}.
If \emph{The-Lord-Of-The-Rings} has no \emph{author} relationship, a constraint violation is raised. 
It leads to invalid data, if \emph{The-Lord-Of-The-Rings} has an \emph{author} property pointing to some resource which is assigned to another class than \emph{Person}
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Human)}).
It also causes a constraint violation, if \emph{The-Lord-Of-The-Rings} has an \emph{author} relation to an object for which no class is associated
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}).
The data is valid, if \emph{The-Lord-Of-The-Rings} is connected to a \emph{person} via the object property \emph{author}
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Person)}).

The following table shows the mapping between the cardinality restriction expressed in DL 
and the cardinality restriction expressed by our developed vocabulary to describe any RDF constraint generically (see section~\ref{sec:ontology}):

\begin{gcotable}
property & Publication & author & - & Person & $\geq$ & 1 \\
\end{gcotable}

According to this vocabulary, each constraint is either a constraint on classes or a constraint on properties.
The generically expressed cardinality restriction is represented in RDF as follows:

\begin{ex}
[   a PropertyConstraint ;
    contextClass Publication ;
    leftProperties ( author ) ;
    classes ( Person ) ;
    constrainingElement ">=" ;
    constrainingValue 1 ] .
\end{ex}

Bosch and Eckert\cite{BoschEckert2014-2} use SPIN as basis to define a
validation environment (available at \url{http://purl.org/net/rdfval-demo}) in which the validation of any constraint language\footnote{the only limitation is that constraint languages must be represented in RDF} can be implemented by representing them in SPARQL. 
The SPIN engine checks for each resource if it satisfies all constraints (associated with its assigned classes) and generates a result RDF graph containing information about all constraint violations.
The generically expressed cardinality restriction above is validated within our validation environment (see section \ref{sec:validation}) by means of a SPARQL CONSTRUCT query:
\begin{ex}
[   a PropertyConstraint ;
    contextClass ?cc ;
    leftProperties ( ?p1 ) ;
    classes ( ?c1 ) ;
    constrainingElement ">=" ;
    constrainingValue ?cv ] .
?subject rdf:type ?cc .
BIND ( qualifiedCardinality( ?subject, ?p1, ?c1 ) AS ?c ) .
FILTER ( ?c < ?cv ) .		  
\end{ex}

\section{A Vocabulary to Describe RDF Constraints Generically} 
\label{sec:ontology}

In order to define a vocabulary to describe RDF constraints generically, it is needed to define the terminology for the formulation of RDF constraints and to classify them. 
A \ms{constraint language} is a language which is used to formulate constraints.
The W3C Data Shapes working group defines \ms{constraint} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
We identified four dimensions to classify constraints:
\begin{itemize}
  \item \ms{Universality:} specific constraints vs. generic constraints
	\item \ms{Complexity:} simple constraints vs. complex constraints
	\item \ms{Context:} property constraints vs. class constraints
	\item \ms{DL Expressivity:} constraints expressible in DL vs. constraints not expressible in DL
\end{itemize}

As there are already five promising constraint languages, our purpose is not to invent a new constraint language.
We rather developed a very simple ontology (only three classes, three object properties, and three data properties) which is universal enough to describe any RDF constraint expressible by any RDF constraint language (see the conceptual model in figure \ref{fig:RDF-CO-conceptual-model}).
We call this vocabulary the \ms{RDF constraints ontology (RDF-CO)}\footnote{Available at: \url{https://github.com/boschthomas/RDF-constraints-ontology}}.

### 2015-02-10 paper/paper7/paper7-appendix.tex
running example

{\em Property Domains}\footnote{{\em R-25-OBJECT-PROPERTY-DOMAIN}, {\em R-26-DATA-PROPERTY-DOMAIN}} ({\em domain restrictions on roles} in DL) restrict the domain of object and data properties.
The purpose is to declare that a given property is associated with a class. 
In OO terms this is the declaration of a member, field, attribute or association. 
$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$.
{\em Property Domain} constraints are defined for each \emph{Disco} object and data property.

\begin{itemize}
	\item \textbf{{\em DISCO-C-PROPERTY-DOMAIN-01:}} 
Only {\em disco:Question}s can have {\em disco:responseDomain} relationships (\ms{$\exists$ responseDomain.$\top$ $\sqsubseteq$ Question}).
\end{itemize}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-PROPERTY-RANGES-01:}} 
{\em disco:caseQuantity} relationships can only point to literals of the datatype {\em xsd:nonNegativeInteger} (\ms{$\top$ $\sqsubseteq$ $\forall$ caseQuantity.nonNegativeInteger}).
\end{itemize}

>>>>

{\em Property Domains}\footnote{{\em R-25-OBJECT-PROPERTY-DOMAIN}, {\em R-26-DATA-PROPERTY-DOMAIN}} ({\em domain restrictions on roles} in DL) restrict the domain of object and data properties.
The purpose is to declare that a given property is associated with a class. 
In OO terms this is the declaration of a member, field, attribute or association. 
$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$.

\begin{itemize}
	\item \textbf{{\em DISCO-C-PROPERTY-DOMAIN-01:}} 
	{\em Property Domain} constraints are defined for each \emph{Disco} object and data property.
  Only {\em disco:Question}s, e.g., can have {\em disco:responseDomain} relationships (\ms{$\exists$ responseDomain.$\top$ $\sqsubseteq$ Question}).
\end{itemize}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-PROPERTY-RANGES-01:}} 
	{\em Property Range} constraints are defined for each \emph{Disco} object and data property.
  {\em disco:caseQuantity} relationships, e.g., can only point to literals of the datatype {\em xsd:nonNegativeInteger} (\ms{$\top$ $\sqsubseteq$ $\forall$ caseQuantity.nonNegativeInteger}).
\end{itemize}
